---
title: "EOS Student Comparison Metrics - FY 13-14 to FY 15-16"
author: "Drew McDermott"
output:  
  flexdashboard::flex_dashboard:
  source_code: embed
runtime: shiny
---

```{r global, include = FALSE}

#==============================================#
# ==== load packages and clear objects/log ====
#==============================================#
  
  library(wesanderson)
  library(data.table)
  library(ggplot2)
  library(plotly)
  library(shiny)
  library(knitr)
  library(flexdashboard)

#====================#
# ==== set parms ====
#====================#
  
  # set working directory to code location
  setwd("C:/Users/Drew/Dropbox/github_clones/personal_code/eos_excercise")

  # set wes anderson color scheme
  wes_pal <- wes_palette(name = "Zissou", type = "continuous")

  # set up base plot attributes / theme 
  plot_attributes <- theme(text       = element_text(size = 20),
                           plot.title = element_text(vjust = 0, colour = "black", face = "bold", size = 25))

#====================#
# ==== load data ====
#====================#
  
  # load data
  load("flex_dash_data.rdata")

  # copy loaded data
  dt_cl_gaps_overall   <- copy(out_list$tb_cl_gaps)
  dt_cl_gaps_srt_yr_wd <- copy(out_list$tb_cl_gaps_srt_yr_wd)
  dt_cl_gaps_srt_yr    <- copy(out_list$tb_cl_gaps_srt_yr)
  dt_st_added          <- copy(out_list$tb_st_added)
  dt_st_added_wide     <- copy(out_list$eos_data_st_added)
  dt_sus_gap           <- copy(out_list$tb_sustain_gap_cl)

```

Closing Gaps {data-orientation=columns}
=====================================  

Column
-----------------------------------------------------------------------

### Chart 1A - Gaps Closed between Underrepresented and Benchmark Students - Number of Schools

```{r, fig.width = 10, fig.height = 7}
  
  # plot - num schools closed gaps, by year
  plot_gaps_yr <- ggplot(data = dt_cl_gaps_overall, aes(x = data_yr, y = value, fill = variable)) +
                    geom_bar(stat     = "identity",
                             position = "identity") +
                    scale_y_continuous(breaks = seq(-100, 100, 25),
                                       limits = c(-100, 100)) +
                    scale_fill_manual(values = c("#3B9AB2", "#F21A00"),
                                      labels = c("Closed Gaps", "Did Not Close Gaps")) +
                    theme(legend.title = element_blank()) +
                    labs(x = "Year (FY)",
                         y = "Number of Schools") +
                    plot_attributes

  # display plot
  plot(plot_gaps_yr)
  
  # output$plot_gaps_yr <- renderPlotly({
  # 
  #   p <- plot_ly(subset(dt_cl_gaps_overall, variable == "closed_gaps"), x = data_yr, y = value,
  #                type = "bar", color = variable)
  #   
  #   p <- plot_ly(dt_cl_gaps_overall, x = data_yr, y = value,
  #              type = "bar", color = variable, marker = list(color = c("#3B9AB2", "#F21A00")))
  #   
  #   p_final <- layout(p, barmode = 'relative')
  # 
  # })
  # 
  # # display plot (as plotly)
  # plotlyOutput('plot_gaps_yr')
  
```

### Chart 1B - Gaps Closed between Underrepresented and Benchmark Students - Percentage of Schools

```{r, fig.width = 10, fig.height = 7}
  
  # plot - perc schools closed gaps, by year
  plot_perc_gaps_yr <- ggplot(data = dt_cl_gaps_overall, aes(x = data_yr, y = percent, fill = variable)) + 
                        geom_bar(stat     = "identity", 
                                 position = "identity") +
                        scale_y_continuous(breaks = seq(-1, 1, .25), 
                                           limits = c(-1, 1)) +
                        scale_fill_manual(values = c("#3B9AB2", "#F21A00"), 
                                          labels = c("Closed Gaps", "Did Not Close Gaps")) +
                        theme(legend.title = element_blank()) +
                        labs(x = "Year (FY)", 
                             y = "Percentage of Schools") +
                        plot_attributes

  # display plot
  plot(plot_perc_gaps_yr)

```

Column
-----------------------------------------------------------------------

### Chart 1C - Closed Gaps Sustained Between Years Gaps

```{r, fig.width = 6, fig.height = 4}
  
  # plot
  plot_sustain_gap_cl <- ggplot(subset(dt_sus_gap, variable == "sustain_gap_cl"), aes(start_year_with_eos, value)) +   
                          geom_bar(aes(fill = data_yr), stat     = "identity",
                                                        position = "dodge") + 
                          geom_label(aes(label = paste0(round(perc_sustain_gap_cl*100, 0), " %"), 
                                         group = data_yr),
                                     position = position_dodge(width = 1)) +
                          scale_fill_manual(values = c("#3B9AB2", "#E1AF00"), 
                            labels = c("Yr 1 to Yr 2", "Yr 2 to Yr 3")) +
                          theme(legend.title = element_blank()) +
                          labs(x = "EOS Start Year",
                               y = "Number of Schools that 
Sustained Gaps Closed") +
                          plot_attributes

  # display plot
  plot(plot_sustain_gap_cl)

```

Closing Gaps (Part 2) {data-orientation=columns}
=====================================  

Column
-----------------------------------------------------------------------

### Table 2A - Summary of Gaps Closed

```{r}

  # rename vars for display
  setnames(dt_cl_gaps_srt_yr_wd, c("EOS Start Year", "Data Year", "Total Schools", "Closed Gaps - Number of Schools", 
                                "Did Not Close Gaps - Number of Schools", "Closed Gaps - Percentage of Schools"))

  # render shiny table
  output$table_summary_gaps <- renderDataTable(dt_cl_gaps_srt_yr_wd)

  # display table
  dataTableOutput('table_summary_gaps')

```


Column
-----------------------------------------------------------------------

### Chart 2B - Number of Schools by EOS Start Year

```{r, fig.width = 12}

  # plot - num schools closed gaps, by year, facetted by start year
  plot_gaps_yr_start_yr <- ggplot(data = dt_cl_gaps_srt_yr, aes(x = data_yr, y = value, fill = variable)) + 
                            geom_bar(stat     = "identity", 
                                     position = "identity") + 
                            facet_wrap("start_year_with_eos") +
                            scale_y_continuous(breaks = seq(-75, 75, 25), 
                                               limits = c(-75, 75)) +
                            scale_fill_manual(values = c("#3B9AB2", "#F21A00"), 
                                              labels = c("Closed Gaps", "Did Not Close Gaps")) +
                            theme(legend.title = element_blank()) +
                            labs(x = "Year (FY)", 
                                 y = "Number of Schools") +
                            plot_attributes

  # display plot
  plot(plot_gaps_yr_start_yr)

```

### Chart 2C - Percentage of Schools by EOS Start Year

```{r, fig.width = 12}

  # plot - perc schools closed gaps, by year, facetted by start year
  plot_perc_gaps_yr_start_yr <- ggplot(data = dt_cl_gaps_srt_yr, aes(x = data_yr, y = percent, fill = variable)) + 
                                  geom_bar(stat     = "identity", 
                                           position = "identity") +
                                  facet_wrap("start_year_with_eos") +
                                  scale_y_continuous(breaks = seq(-1, 1, .25), 
                                                     limits = c(-1, 1)) +
                                  scale_fill_manual(values = c("#3B9AB2", "#F21A00"), 
                                                    labels = c("Closed Gaps", "Did Not Close Gaps")) +
                                  theme(legend.title = element_blank()) +
                                  labs(x = "Year (FY)", 
                                       y = "Percentage of Schools") +
                                  plot_attributes

  # display plot
  plot(plot_perc_gaps_yr_start_yr)

```

Underrepresented Students Added {data-orientation=columns}
=====================================

Inputs {.sidebar}
-----------------------------------------------------------------------

```{r}

  # define checkboxes for cohort selection
  checkboxGroupInput(inputId = "selected_cohort_p2", 
                     label = h3("EOS Start Year (FY):"),
                     choices = c("FY13-FY14", "FY14-FY15", "FY15-FY16"),
                     selected = c("FY13-FY14", "FY14-FY15", "FY15-FY16"))

  # define checkboxes for data year selection
  checkboxGroupInput(inputId ="selected_yr_p2", 
                     label = h3("Data Year (FY):"),
                     choices = c("2014", "2015", "2016"),
                     selected = c("2014", "2015", "2016"))
  
  
  # combine the selected variables into a new data frame - box and whisker
  selected_data_p3a <- reactive({ output <- subset(dt_st_added, chmatch(`start_year_with_eos`, c(input$selected_cohort_p2), nomatch = 0) != 0 & 
                                                       chmatch(`data_yr`, c(input$selected_yr_p2), nomatch = 0) != 0) })

  # combine the selected variables into a new data frame - box and whisker
  selected_data_p3b <- reactive({ output <- subset(dt_st_added_wide, num_ap_students_added > 20 & 
                                                       chmatch(start_year_with_eos, c(input$selected_cohort_p2), nomatch = 0) != 0) })
  
  # combine the selected variables into a new data frame - histogram
  selected_data_p3c <- reactive({ output <- subset(dt_st_added_wide, num_ap_students_added > 20 & 
                                                       chmatch(start_year_with_eos, c(input$selected_cohort_p2), nomatch = 0) != 0 & 
                                                       chmatch(data_yr, c(input$selected_yr_p2), nomatch = 0) != 0) })

```

Column
-----------------------------------------------------------------------

### Chart 3B - Distribution of Percentage Underrepresented Students Added to AP/IB Over Baseline, by Year

```{r, fig.width = 7}

  # render output chart 2A
  output$plot_box_ur_added <- renderPlotly({

    # plot - box and whisker of percent of ur students added (total students added > 20)
    p3b <- ggplot(data = selected_data_p3b(), aes(x = data_yr, y = perc_ap_ur_students_added)) +
            geom_boxplot(colour         = "black", 
                         fill           = "#3B9AB2") +
            stat_summary(fun.y = mean, 
                         geom  = "point", 
                         shape = 5, size = 3) +
            labs(x = "Year (FY)", 
                 y = "Percent Underrepresented Students")
    
    # add plotly attributes
    ggplotly(p3b) %>%
      layout(xaxis = list(titlefont = list(size = 18)), 
             yaxis = list(titlefont = list(size = 16)))

  })

  # display plot (as plotly)
  plotlyOutput('plot_box_ur_added')


```


### Chart 3C - Distribution of Percentage Underrepresented Students Added to AP/IB Over Baseline

```{r, fig.width = 7}

  # render output chart 2B
  output$plot_hist_ur_added <- renderPlotly({

    # plot - histogram of percent of ur students added (total students added > 20)
    p3c <- ggplot(data = selected_data_p3c(), aes(x = perc_ap_ur_students_added)) +
            geom_histogram(binwidth = .05,
                           colour   = "black", 
                           fill     = "#3B9AB2") +
            labs(x = "Percent Underrepresented Students", y = "Number of Schools")
    
    # add plotly attributes
    ggplotly(p3c) %>%
      layout(xaxis = list(titlefont = list(size = 18)), 
             yaxis = list(titlefont = list(size = 16)))

  })

  # display plot (as plotly)
  plotlyOutput('plot_hist_ur_added')

```

Column
-----------------------------------------------------------------------

### Table 3A - Summary of Students Added

```{r}

  selected_data_p3a_rename <- reactive({ setnames(selected_data_p3a(), 
                                                  c("EOS Start Year", "Data Year", "Total Schools", "Avg. Students Added to AP/IB Over Baseline", 
                                                    "Avg. Underrepresented Students Added", "Avg. Benchmark Students Added", 
                                                    "Avg. Percent Underrepresented")) })

  # render shiny table
  output$table_st_add_avg <- renderDataTable(selected_data_p3a_rename())

  # display table
  dataTableOutput('table_st_add_avg')

```


Percentage of Underrepresented Students Added by School
=====================================

Inputs {.sidebar}
-----------------------------------------------------------------------

Select the fiscal year (FY) to display in the chart to the right

```{r}

  # define checkboxes for data year selection
  checkboxGroupInput(inputId ="selected_yr_p3", 
                     label = h3("Data Year (FY):"),
                     choices = c("2014", "2015", "2016"),
                     selected = c("2014", "2015", "2016"))

  # combine the selected variables into a new data frame
  selected_data_p4 <- reactive({ 
    
    # subset to schools with at least 20 students added, and selected data year
    p4 <- subset(dt_st_added_wide, num_ap_students_added > 20 & chmatch(data_yr, c(input$selected_yr_p3), nomatch = 0) != 0)
    
    # rename variables for display
    setnames(p4, c("num_ap_ur_students_added", "perc_ap_ur_students_added", "start_year_with_eos"), 
             c("Students Added", "Percent UR", "EOS Start Year"))
      
  })

```

Outputs
-----------------------------------------------------------------------

### Chart 4A - Scatter Plot of Underrepresented Students Added to AP/IB Over Baseline

```{r}

  # render output chart 3A
  output$plot_scatter_ur_added <- renderPlotly({

    # plot - scatter of schools, students added by perc ur students added (total students added > 20)
    p4a <- ggplot(data = selected_data_p4(), aes(x = `Students Added`, y = `Percent UR`)) + 
                            geom_point(aes(color = `EOS Start Year`,
                                           text = paste("ID: ", district_id, "-", school_id)),
                                       size   = 5) +
                            labs(x = "Total Number of Students Added to AP/IB Over Baseline", 
                                 y = "Percent Underrepresented Students") +
                            plot_attributes
    
    # add plotly attributes
    ggplotly(p4a) %>%
      layout(xaxis = list(titlefont = list(size = 18)), 
             yaxis = list(titlefont = list(size = 15)),
             margin = list(r = 300))

  })

  # display plot (as plotly)
  plotlyOutput('plot_scatter_ur_added')

```

